\chapter[James McDonald]{\mcdonaldJName{}}
\label{ch1: James McDonald}

\genealogyInfoBox[a]{\male~\mcdonaldJName{}}{
    \bornEntry{\DTMDate{1789-10-9}}{\crawfordFull{}, \scotland{}} \\
    \christenedEntry{\DTMDate{1789-10-11}}{\scotland{}} \\
    \diedEntry{\before~\DTMDate{1822-11-07}}{\capeColony, \southAfrica}
}{
    \marriedEntry{\DTMDate{1817-05-12}}{\stBotolphWithoutBishopsgate{}}{\welchMName}
}

\section{Early life in Britain}

\mcdonaldJNameOnly{}'s birth date, birthplace, and parents remain uncertain. The most reliable source is \biggarASurname{}'s \settler{} list, which listed his age as 31 \autocite[536-539]{eggsa:biggarCorrespondance}. This suggests a birth year between \DTMdate{1787-09-28}, and \DTMdate{1788-09-27}. However, settlers often inflated their ages to qualify for the program, so this estimate should be viewed with caution.

\citeauthor{FS:JamesMcDonaldBaptism} and \citeauthor{settlers:JamesMcDonald} suggest \mcdonaldJName{} was born on \DTMdate{1789-10-09}, to \mcdonaldWWFullNames{} and \aitchisonJFullNames{} in \crawfordFull{}, \scotland{}, with a baptism on \DTMdate{1789-10-11} \autocite{FS:JamesMcDonaldBaptism} \autocite{settlers:JamesMcDonald}. However, this lacks evidence to definitively link them. The birth date contradicts the age calculation above, and the suggested parents' names haven't reappeared among known descendants.

\section[The Journey to the Cape Colony]{The Journey to the \capeColony}
\label{ch1: Journey to the Cape Colony}

\mcdonaldJNameOnly{}, \welchMNameOnly{}, and their one-year-old son, \mcdonaldANameOnly{}, immigrated to \southAfrica{} with the \settlersBritish{} \autocite[46]{nash:1820}. Their names first appear in \biggarAName{}'s settler correspondence. This letter, dated \DTMdate{1819-09-27}, was addressed to \bathhurstHName{} from 3 \northumberlandCourtFull{}, \london{}, in which \biggarASurname{} proposed a list of potential members for his party to \southAfrica \autocite[536-539]{eggsa:biggarCorrespondance}. They are also referenced in his later settler returns (Appendix 1.3.3 provides a complete list and further details) \autocite{eggsa:biggarReturn}.

\section[{Settlement and Life in the Cape Colony}]{Settlement and Life in the \capeColony}

\section{Marriage and Family}

\subsection[Marriage to Mary Welch]{Marriage to \welchMName}
\label{ch1: Marriage to Mary Welch}

\mcdonaldJName{} and \welchMName{} were wed on the \DTMdate{1817-05-12}, at \stBotolphWithoutBishopsgate{}, \bishopsgateFull{} \autocite{FS:JamesMcDonaldMarriage} \autocite{settlers:JamesMcDonald}. The marriage record shows \mcdonaldJNameOnly{} signing his name, while \welchMNameOnly{} made a mark ("+") next to hers. Both were listed as parishioners. \nettoIName{} and \nettoCNameNee{} witnessed the ceremony (Figure 1.1).

\section{Descendants}

\mcdonaldAName{}, born from the union of \mcdonaldJName{} and \welchMName{}, is their only known descendant \autocite[46]{nash:1820}.

\section{Death}

\mcdonaldJNameOnly{} must have died before the \DTMdate{1822-11-07} because we know that marriage bans were published on that date for \welchMName{} and \henmanRName{}. See Chapter \ref{ch2: Marriage to Robert Henman} for further details.

\section{Associates}